---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
------------------------------------------


Author:  C. Deramond  
Course:  *repdata-008* 


## Loading and preprocessing the data

For this assignment, we initially load packages, data and make initial transformations:
```{r}
library(ggplot2, quietly = T, warn.conflicts = F)
library(lubridate, quietly = T, warn.conflicts = F)
library(dplyr, quietly = T, warn.conflicts = F)

data <- read.csv("activity.csv", stringsAsFactors=FALSE) #get the data
data$date <- ymd(data$date) # date parsing              

```


## What is mean total number of steps taken per day?

let's examine code to generate response first:

```{r}
#summary table
data <- as.tbl(data) %>%
        filter(!is.na(steps)) %>%
        group_by(date) %>%
        summarise(n = n(), total = sum(steps), mean = mean(steps), median = median(as.numeric(steps))) 
```
1. **histogram:**
```{r}
hist <- ggplot(data)
hist + aes(total)  + 
        geom_histogram(fill="steelblue") + 
        ggtitle("Histogram: total steps by day") + 
        labs(x = "Total steps")

```

2. **summary table:** 
```{r echo=FALSE}
data
```




## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
